### 2023-12-01 paper/content.tex
Add a working rule for making diffs
Finished paper review after R1

Several branches of Earth Sciences have demonstrated the importance of spatial resolution on a microscopic scale. For example, geochemistry and
geochronology applications have benefited from point-wise analyses
and compositional maps, allowing significant advances in the understanding of
igneous, metamorphic and sedimentary processes \citep{Verberne2020, Barnes2019,
Davidson2007}. Classical paleomagnetic techniques, on the other hand, consist
of analyzing bulk samples, where the magnetic signal of a single specimen is
the result of the sum of moments of a large assembly of ferromagnetic grains
\citep{Dunlop1997}. Typically, a standard paleomagnetic sample of
approximately \qty{10}{\cm\cubed} would contain hundreds of thousands
to millions of magnetic particles with sizes varying from magnetically stable
single-domain (SD) and vortex state grains (also called pseudo-single domain,
PSD) with sizes below \qty{1}{\um} \citep{Berndt2016}, to large ($\gg \qty{1}{\um}$) grains with
multi-domain (MD) magnetic structures, which are less stable magnetic recorders \citep{Dunlop1997}.
These large MD grains usually conceal the signal of the SD
and PSD grains, and techniques of step-wise thermal and magnetic treatments are
needed to unveil this more stable and reliable magnetic record
\citep{Tauxe2018}. Recently, magnetic microscopy techniques opened the
possibility of obtaining magnetic field maps at the micro-scale and recovering the
magnetization of each grain, therefore enabling the separate analysis of stable
and unstable magnetic particles \citep{DeGroot2018, Lima2014, Weiss2007,
DeGroot2014}.

In order to apply magnetic microscopy to paleomagnetic studies, it is necessary
to recover from the magnetic images a large number of individual magnetic
moments, corresponding to at least tens of thousands of stable fine-grained
grains ($< \qty{1}{\um}$), in order to provide statistical significance to the
remanence vector \citep[e.g., ][]{Berndt2016}. Nowadays, with the development of
magnetic microscopy techniques, this task is no longer limited by the
resolution of magnetic microscopes \citep{Fu2020, Weiss2007, DeGroot2018,
Glenn2017, Lima2014}, but essentially by the intrinsic problem presented by the
ambiguity in the inversion of potential field data \citep{Barbosa2011,
DeGroot2021, Oliveira2015Estimation}, and ultimately by the lack of a fast and
automated way to recover such a large number of individual magnetic moments
from a set of magnetic images \citep{CortesOrtuno2022, Lima2013, Lima2009}. A
solution to the non-uniqueness of magnetic moment inversion is to add
independent prior information, such as the position of the ferromagnetic
particles \citep{Fabian2019}. This can be obtained, for example, from X-ray
computed tomography \citep[microCT; ][]{Fabian2019, DeGroot2021, DeGroot2018}.
Nonetheless, the standard microCT techniques do not provide adequate resolution
to resolve the finer and more stable magnetic grains \citep{CortesOrtuno2022,
DeGroot2021}, whereas other more sophisticated techniques such as ptychographic
X-ray tomography \citep[e.g., ][]{Maldanis2020} are not readily available and
too time-consuming to be routinely used in paleomagnetic studies.

Another route to be explored in the inversion of magnetic microscopy images is
to obtain all the information, i.e. the magnetic moment and the position of the
sources, from the magnetic data itself \citep[e.g., ][]{Fu2020}. For that, we can
explore the techniques developed in exploration geophysics, in spite of the
differences between aeromagnetic surveys and magnetic microscopy
\citep{Lima2013}. Magnetic microscopy images commonly show the combined signal
of multiple magnetic particles and can vary greatly in wavelength, strength,
and spatial separation, depending on the natural remanent magnetization (NRM) and location of each particle. We
usually assume that the signal measured by the magnetic microscope is the
vertical component of the magnetic induction vector ($b_z$), the measurements
are performed on a regular grid with evenly spaced grid points and at a
constant height, and the data are contaminated with pseudorandom Gaussian noise and
long-wavelength noise (akin to a regional signal in aeromagnetic data). Here,
we provide a methodological routine to retrieve the individual magnetic moment
of ferromagnetic grains in magnetic microscopy images following the approach
devised by \citet{Oliveira2015Estimation} for the interpretation of
aeromagnetic anomalies. The method we propose allows one to quickly and
semi-automatically estimate the individual magnetic moment vector of the stable
magnetic carriers, making use of only the magnetic images themselves and an
assumption of approximately dipolar sources. If used on a large scale, the
method provides the means to scan large areas of the rock sample, attaining
potentially the number of magnetic moments necessary for paleomagnetic studies.

@@

\begin{enumerate}
\item \textbf{Source detection and separation:} Identify and spatially isolate
  the magnetic field caused by each source. We will do this through a
  combination of classic potential field data processing (\textit{total
  gradient amplitude}) and image processing (histogram stretching and
  equalization) and segmentation methods (watershed segmentation).
\item \textbf{Position estimation}: Estimate the 3D position of a magnetic
  particle based on the magnetic field measurements and the assumption of a
  dipolar source. This can be achieved by applying the Euler Deconvolution
  method to the data segment identified in step 1.
\item \textbf{Magnetic moment inversion:} Estimate the 3-component dipole
  moment vector by inverting the magnetic field data using the position
  obtained from Euler Deconvolution as a constraint and the assumption of a
  dipolar source. This leads to a linear inverse problem that is stable and
  computationally efficient, particularly since the inversion is performed
  separately for each data segment identified in step 1.
\end{enumerate}

@@

\noindent
and likewise for the $y$ and $z$ directions, in which $\Delta x$ is the grid
spacing (assumed to be equal in the $x$ and $y$ directions). For the $z$
component, $\Delta z = \Delta x$ and $f(x, y, z + \Delta z)$ and
$f(x, y, z - \Delta z)$ are calculated by upward and downward continuation,
respectively, performed in the wavenumber domain. The 2D maps of the $x$, $y$,
and $z$ derivatives will also be used in the Euler Deconvolution step described
below, which is known to be highly sensitive to noise in the derivatives
\citep{Saleh2012Applying}. This is why we prefer the finite-differences based
derivatives, which are less prone to amplifying short-wavelength noise than
those calculated in the wavenumber domain through the Fast Fourier Transform
(FFT).

@@

Once the locations and sizes of the windows containing the isolated signals of
each particle are determined, we can apply the Euler Deconvolution (ED)
estimate of the $x$, $y$, and $z$ positions of the center of the particles
under a dipolar approximation. This technique was first proposed by
\citet{Thompson1982} under the name EULDPH and later extended to three
dimensions and renamed ``Euler Deconvolution'' by \citet{Reid1990}. ED is
traditionally performed on a set of moving windows that scan the entire
dataset, producing a large scatter of position estimates, most of which are
spurious \citep{Silva20033D}. This is only done because it is impractical to
segment an aeromagnetic dataset into individual anomalies. Fortunately,
magnetic microscopy images often contain fewer elongated features (e.g., dikes and
suture zones) and regional signals (e.g., Curie depth variations) that are
difficult to separate. This makes it possible for us to produce isolated
signals for each magnetic particle through the source detection step described
above. Since each data window contains only the signal of a single particle, we
are able to apply ED and generate only a single position estimate per particle.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/coordinate-system-sketch.png}
\caption{
Schematic representation of the coordinates systems and modelling elements.
The left panel shows the $x, y, z$ right-handed coordinate system with $z$
pointing upward and away from the sample. The top surface of the sample defines
the $z=0$ surface. A dipole with dipole moment vector $\mathbf{\vec{m}}$ is
also shown at coordinates $(x_c, y_c, z_c)$. The observation points are located
at regular intervals on an $x-y$ plane at a positive $z$ separation from the
sample surface. The right panel zooms in on the dipole and shows the dipole
moment vector expressed in terms of inclination ($I$, positive downwards),
declination ($D$, angle with the $\hat{y}$ direction), and moment magnitude $m
= \|\mathbf{m}\|$.
}
\label{fig_coordinate_systems}
\end{figure}

@@

Once the source position is known and we can assume that it is approximately
spherical or dipolar, we can apply the method developed by
\citet{Oliveira2015Estimation} to estimate the dipole moment vector
$\mathbf{m}$ of the source. We begin by following
\citet{Oliveira2015Estimation} in formulating the magnetic induction vector
$\mathbf{b}$ of a dipole as

@@

\noindent
From the main diagonal of $\mathbf{C}$, we can obtain the variances of the
estimated declination $\sigma_D^2$, inclination $\sigma_I^2$, and magnitude $\sigma_m^2$
by propagation of uncertainties from Equation~\ref{eq_vector_to_incdec}

@@

We first applied our inversion workflow to three sets of synthetic data to evaluate its strengths and limitations.
The datasets are produced by different models of varying complexity which were designed to assess different aspects of the method.

@@

\item \textbf{Applicability to real-world scenarios:}
The third model contains 103 dipoles with different dipole moment magnitudes, having inclinations and declinations clustered around two stable directions.
The magnetic field generated from these sources is corrupted by both low and high-frequency noise.
The complexity of this synthetic data seeks to more faithfully emulate real magnetic microscopy data.
\end{enumerate}

@@

The positions of each source obtained with Euler deconvolution were then used as input for the dipole moment inversion.
The estimated dipole moment vectors are shown in Figure~\ref{fig_synthetic_simple_results}.
Figure~\ref{fig_synthetic_simple_results}a shows the estimated dipole moment, along with the corresponding true values, as a stereonet for better comparison of the true and estimated vector directions.
Figure~\ref{fig_synthetic_simple_results}b shows the estimate dipole moments overlaid on a map of the synthetic $b_z$ to demonstrate the ability of our method to estimate a spatial distribution of dipole moments.
For comparison, Table~\ref{tab_synthetic_simple_results} also shows the true and estimate dipole moments as well as their corresponding standard deviations obtained using Equations~\ref{eq_chi_square}-\ref{eq_variances}.

@@

To simulate particles with non-dipolar magnetization components, we specified a spherical volume with a radius of approximately \qty{1}{\um}.
Within this volume, we added 200 dipolar particles with varying dipole moment directions and amplitudes.
These directions were generated randomly, following a normal distribution centered on the direction $D=\ang{90}$ and $I=\ang{0}$ and with a standard deviation of \ang{10}.
The dipole moment amplitudes were sampled by a normal distribution with mean $\qty{e-14}{\ampere\m\squared}$ and standard deviation $\qty{5e-14}{\ampere\m\squared}$.
We also randomly generated spatial positions around a central position of $x_c=\qty{25}{\um}$, $y_c=\qty{25}{\um}$, and $z_c=\qty{-1}{\um}$ with standard deviation of \qty{0.65}{\um}.
The bulk magnetization of the non-dipolar particle is the vector sum of the dipole moments of each dipole.
The centroid of the spherical volume is defined by the average of the $x_c$, $y_c$, and $z_c$ positions of each dipole.
The synthetic vertical magnetic field $b_z$ of the non-dipolar source was calculated on regular grids with \qty{0.5}{\um} spacing.
Each grid was located at a different sensor-sample distance $z$, varying between \qty{1}{\um} and \qty{10}{\um} in \qty{0.5}{\um} increments.
The synthetic data were corrupted by pseudo-random Gaussian noise of mean \qty{0}{\nano\tesla} and a standard deviation of \qty{20}{\nano\tesla}.
For each data grid generated, we performed Euler deconvolution to estimate the source position and subsequently inverted the data for the dipole moment.
At each step, we recorded the inversion $R^2$ value and the differences between the estimated position and dipole moment and the true centroid and bulk magnetization of the spherical particle.

@@

We replicated the particle generation procedure 100 times with variations in the randomness of magnetization and point particle positions. Subsequently, the Cartesian position of the spherical particle was estimated using ED, and the magnetization parameters were determined by inverting the magnetic field data. The effectiveness of position recovery was measured by comparing the centroid of the spherical volume with the solution of the Euler equation for $x_c$, $y_c$, and $z_c$ (Figure~\ref{non-dipolarity-synthetic-data-positioning}a, b, and c, respectively). As shown in Figure~\ref{non-dipolarity-synthetic-data-positioning}, ED estimates the positions of particles with more complex magnetization well, especially for sensor-source distances greater than 5 µm, where the median of the distributions is centered at zero. However, when the sensor-source distance is small enough to emphasize the contribution from higher-order components, a lower degree of dipolarity is observed as expressed by lower values of $R^2$ (Figure~\ref{non-dipolarity-synthetic-data-inversion}a). Consequently, in these cases, there are higher errors in the estimation of the direction of magnetization (Figure~\ref{non-dipolarity-synthetic-data-inversion}b) and in the magnetic moment's intensity (Figure~\ref{non-dipolarity-synthetic-data-inversion}c), which were expected since the inversion considers only the dipolar component. However, dipolarity significantly increases as the sensor to sample distance increases, while errors in direction and magnetic moment decrease significantly (Figure~\ref{non-dipolarity-synthetic-data-inversion}).

@@

For greater fidelity to real samples, the magnetic sources are modeled as dipoles with different depths and magnetic moment intensities.
The depths vary randomly between 1 and $\qty{20}{\um}$, while the dipole moment intensities range randomly from $10^{-11}$ to $\qty{e-14}{\ampere\m\squared}$.
The NRM found in real ferromagnetic particles varies individually but averages out to the inducing field direction.
To simulate this behavior in our synthetic data, we sample the source dipole moment directions from two pseudo-random Gaussian distributions.
The first group of sources ($M = 70$) are sampled from a distribution  with mean of $D = \ang{0}$ and $I = \ang{0}$ and standard deviation of $\ang{10}$.
The second group of sources ($M = 30$) are sampled from a distribution with mean of $D = \ang{180}$ and $I = \ang{0}$, also with standard deviation of $\ang{10}$. We also manually added 3 sources with higher dipole moments ($5$x$10^{-11}$) to further simulate the complexity observed in real data measurements. The noise-corrupted synthetic $b_z$ data are shown in Figure~\ref{complex-synthetic-data}a.

@@

We then followed the same processing steps as for the simple synthetic: upward continuation (Figure~\ref{complex-synthetic-data}b),
TGA calculation (Figure~\ref{complex-synthetic-data}c), contrast stretching (Figure~\ref{complex-synthetic-data}d), blob detection (Figure~\ref{complex-synthetic-data}e), and Euler Deconvolution (Figure~\ref{complex-synthetic-data}e).
A total of 99 sources out of the original 103 were successfully detected in approximately 8 seconds by our algorithm. It further determined the 3D positions and dipole moments of these sources in just 1.5 seconds, all while running on an Intel(R) Core(TM) i7-2670QM CPU 2.20GHz with 16 GB of DDR3L 1333 MHz RAM. Four sources remained unidentified (black stars, Figure~\ref{complex-synthetic-data}f) due to specific challenges, that arose from two primary factors. First, three of these sources either were overlapped by or were in close proximity to other neighboring sources, making it inherently difficult to distinguish and isolate them. Second, the identification of one of these sources was impeded by its considerable depth, which led to substantial signal attenuation, further exacerbating the difficulties associated with source recognition.

\begin{figure}[tb!]
\centering
\includegraphics[width=0.75\linewidth]{figures/complex-synthetic-comparison.png}
\caption{
The validation of the result obtained with the inversion was calculated for each individual particle based on the error between the real parameters modeled and their respective recovered values, being (a) the direction and (b) intensity of the magnetic moment, in addition to the $R^2$ score (c) obtained by comparing the forward model and the actual data. The depth and radius of the magnetic sources are also important factors that influence the final result, therefore, these data are given in the form of cross plots, with the vertical bar represented by the depth (1 - 20 $\mu m$) and the horizontal bar by the dipole moment amplitude ($10^{-11}$ to $10^{-14}$ $Am^2$).
}
\label{complex-synthetic-comparison}
\end{figure}

The spatial resolution of the inversion results is one of the key advantages of magnetic microscopy over the classic techniques of paleomagnetism.
Therefore, we present the inversion results spatially so that we can evaluate any patterns in their distribution.
Figure~\ref{complex-synthetic-comparison} shows the spatial locations of the 99 sources that were identified and the differences between the estimated dipole moments and the true values.
The depth and dipole moment amplitude of each source are represented by horizontal and vertical bars, respectively.
These two variables are useful proxies for the strength and spatial extent of the signal of each source,
which can tell us about the limitations in strengths and sizes of the source signal that the technique is able to correctly invert.
Figure~\ref{complex-synthetic-comparison}a shows the absolute value of the angular difference between the true and the estimated dipole moment vectors.
Figure~\ref{complex-synthetic-comparison}b shows the percentage difference between the true and estimated dipole moment magnitudes ($|100 \left( m_{true} - m_{estimated}\right) ~/~ m_{true}|$).
Figure~\ref{complex-synthetic-comparison}c shows the $R^2$ coefficient (Equation~\ref{eq_r2}), which is equivalent to the non-dipolarity parameter of \citet{Fu2020} and represents how well the dipolar model is able to explain the observed data of each source.
Figure~\ref{complex-synthetic-comparison}d shows the SNR (Equation~\ref{eq_snr}), which expresses the power of the observed data over that of the inversion residuals.
High SNR values correspond to small inversion residuals which indicate that a dipolar model was able to explain the observed data.
Consequently, the variation of SNR values is similar to that of the $R^2$ coefficient.
Figure~\ref{complex-synthetic-comparison} shows that large errors in the estimated dipole moment direction are associated with low values of $R^2$ and SNR.
Conversely, the correlation between errors in dipole moment magnitude and $R^2$ and SNR is less pronounced, with some poor magnitude estimates being associated with $R^2$ and SNR indicating a reasonable fit by dipolar model.
It is also noticeable that the majority of cases where the direction misfit is given by deep and low-amplitude sources that are close to shallower or higher-amplitude sources.
These results indicated that $R^2$ and SNR can be used as selection criteria to discard sources with likely high errors in the estimated dipole moment.

>>>>

Several branches of Earth Sciences have demonstrated the importance of spatial resolution on a microscopic scale. For example, geochemistry and
geochronology applications have benefited from point-wise analyses
and compositional maps, allowing significant advances in the understanding of
igneous, metamorphic and sedimentary processes \citep{Verberne2020, Barnes2019,
Davidson2007}. Classical paleomagnetic techniques, on the other hand, consist
of analyzing bulk samples, where the magnetic signal of a single specimen is
the result of the sum of moments of a large assembly of ferromagnetic grains
\citep{Dunlop1997}. Typically, a standard paleomagnetic sample of
approximately \qty{10}{\cm\cubed} would contain hundreds of thousands
to millions of magnetic particles with sizes varying from magnetically stable
single-domain (SD) and vortex state grains (also called pseudo-single domain,
PSD) with sizes below \qty{1}{\um} \citep{Berndt2016}, to large ($\gg \qty{1}{\um}$) grains with
multi-domain (MD) magnetic structures, which are less stable magnetic recorders \citep{Dunlop1997}.
These large MD grains usually conceal the signal of the SD
and PSD grains and techniques of step-wise thermal and magnetic treatments are
needed to unveil this more stable and reliable magnetic record
\citep{Tauxe2018}. Recently, magnetic microscopy techniques opened the
possibility of obtaining magnetic field maps at the micro-scale and recovering the
magnetization of each grain, therefore enabling the separate analysis of stable
and unstable magnetic particles \citep{DeGroot2018, Lima2014, Weiss2007,
DeGroot2014}.

In order to apply magnetic microscopy to paleomagnetic studies, it is necessary
to recover from the magnetic images a large number of individual magnetic
moments, corresponding to at least tens of thousands of stable fine-grained
grains ($< \qty{1}{\um}$), in order to provide statistical significance to the
remanence vector \citep{Berndt2016}. Nowadays, with the development of
magnetic microscopy techniques, this task is no longer limited by the
resolution of magnetic microscopes \citep[\textit{e.g.}, ][]{Fu2020, Weiss2007, DeGroot2018,
Glenn2017, Lima2014, Reis2016}, but essentially by the intrinsic problem presented by the
ambiguity in the inversion of potential field data \citep{Barbosa2011,
DeGroot2021, Oliveira2015Estimation}, and ultimately by the lack of a fast and
automated way to recover such a large number of individual magnetic moments
from a set of magnetic images \citep{CortesOrtuno2022, Lima2013, Lima2009}. A
solution to the non-uniqueness of magnetic moment inversion is to add
independent prior information, such as the position of the ferromagnetic
particles \citep{Fabian2019}. This can be obtained, for example, from X-ray
computed tomography \citep[microCT; ][]{Fabian2019, DeGroot2021, DeGroot2018}.
Nonetheless, the standard microCT techniques do not provide adequate resolution
to resolve the finer and more stable magnetic grains \citep{CortesOrtuno2022,
DeGroot2021}, whereas other more sophisticated techniques such as ptychographic
X-ray tomography \citep[\textit{e.g.}, ][]{Maldanis2020} are not readily available and
too time-consuming to be routinely used in paleomagnetic studies.

Another route to be explored in the inversion of magnetic microscopy images is
to obtain all the information, \textit{i.e.} the magnetic moment and the position of the
sources, from the magnetic data itself \citep[\textit{e.g.}, ][]{Fu2020}. For that, we can
explore the techniques developed in exploration geophysics, in spite of the
differences between aeromagnetic surveys and magnetic microscopy
\citep{Lima2013}. Magnetic microscopy images commonly show the combined signal
of multiple magnetic particles and can vary greatly in wavelength, strength,
and spatial separation, depending on the natural remanent magnetization (NRM) and location of each particle. We
usually assume that the signal measured by the magnetic microscope is the
vertical component of the magnetic induction vector ($b_z$), the measurements
are performed on a regular grid with evenly spaced grid points and at a
constant height, and the data are contaminated with pseudorandom Gaussian noise and
long-wavelength noise (akin to a regional signal in aeromagnetic data). Here,
we provide a methodological routine to retrieve the individual magnetic moment
of ferromagnetic grains in magnetic microscopy images following the approach
devised by \citet{Oliveira2015Estimation} for the interpretation of
aeromagnetic anomalies. The method we propose allows one to quickly and
semi-automatically estimate the individual magnetic moment vector of the stable
magnetic carriers, making use of only the magnetic images themselves and an
assumption of approximately dipolar sources. If used on a large scale, the
method provides the means to scan large areas of the rock sample, attaining
potentially the number of magnetic moments necessary for paleomagnetic studies.

@@

\begin{enumerate}
\item \textbf{Source detection and separation:} Identify and spatially isolate
  the magnetic field caused by each source. We will do this through a
  combination of classic potential field data processing (\textit{total
  gradient amplitude}) and image processing (histogram stretching and
  equalization) and segmentation methods (watershed segmentation).
\item \textbf{Position estimation}: Estimate the 3D position of a magnetic
  particle based on the magnetic field measurements and the assumption of a
  dipolar source. This can be achieved by applying the Euler deconvolution
  method to the data segment identified in step 1.
\item \textbf{Magnetic moment inversion:} Estimate the 3-component dipole
  moment vector by inverting the magnetic field data using the position
  obtained from Euler Deconvolution as a constraint and the assumption of a
  dipolar source. This leads to a linear inverse problem that is stable and
  computationally efficient, particularly since the inversion is performed
  separately for each data segment identified in step 1.
\end{enumerate}

@@

\noindent
and likewise for the $y$ and $z$ directions, in which $\Delta x$ is the grid
spacing (assumed to be equal in the $x$ and $y$ directions). For the $z$
component, $\Delta z = \Delta x$ and $f(x, y, z + \Delta z)$ and
$f(x, y, z - \Delta z)$ are calculated by upward and downward continuation,
respectively, performed in the wavenumber domain. The 2D maps of the $x$, $y$,
and $z$ derivatives will also be used in the Euler Deconvolution step described
below, which is known to be highly sensitive to noise in the derivatives
\citep{Saleh2012Applying}. This is why we prefer the finite-differences-based
derivatives, which are less prone to amplifying short-wavelength noise than
those calculated in the wavenumber domain through the Fast Fourier Transform
(FFT).

@@

Once the locations and sizes of the windows containing the isolated signals of
each particle are determined, we can apply the Euler Deconvolution (ED)
estimate of the $x$, $y$, and $z$ positions of the center of the particles
under a dipolar approximation. This technique was first proposed by
\citet{Thompson1982} under the name EULDPH and later extended to three
dimensions and renamed ``Euler Deconvolution'' by \citet{Reid1990}. ED is
traditionally performed on a set of moving windows that scan the entire
dataset, producing a large scatter of position estimates, most of which are
spurious \citep{Silva20033D}. This is only done because it is impractical to
segment an aeromagnetic dataset into individual anomalies. Fortunately,
magnetic microscopy images often contain fewer elongated features (\textit{e.g.}, dikes and
suture zones) and regional signals (\textit{e.g.}, Curie depth variations) that are
difficult to separate. This makes it possible for us to produce isolated
signals for each magnetic particle through the source detection step described
above. Since each data window contains only the signal of a single particle, we
are able to apply ED and generate only a single position estimate per particle.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/coordinate-system-sketch.png}
\caption{
Schematic representation of the coordinates systems and modeling elements.
The left panel shows the $x, y, z$ right-handed coordinate system with $z$
pointing upward and away from the sample. The top surface of the sample defines
the $z=0$ surface. A dipole with dipole moment vector $\mathbf{\vec{m}}$ is
also shown at coordinates $(x_c, y_c, z_c)$. The observation points are located
at regular intervals on an $x-y$ plane at a positive $z$ separation from the
sample surface. The right panel zooms in on the dipole and shows the dipole
moment vector expressed in terms of inclination ($I$, positive downwards),
declination ($D$, angle with the $\hat{y}$ direction), and moment magnitude $m
= \|\mathbf{m}\|$.
}
\label{fig_coordinate_systems}
\end{figure}

@@

Once the source's position is known, and assuming it is a dipole, we can apply the method developed by
\citet{Oliveira2015Estimation} to estimate the dipole moment vector
$\mathbf{m}$ of the source. We begin by following
\citet{Oliveira2015Estimation} in formulating the magnetic induction vector
$\mathbf{b}$ of a dipole as

@@

\noindent
From the main diagonal of $\mathbf{C}$, we can obtain the variances of the
estimated declination $\sigma_D^2$, inclination $\sigma_I^2$, and magnitude $\sigma_m^2$
by the propagation of uncertainties from Equation~\ref{eq_vector_to_incdec}

@@

We first applied our inversion workflow to four sets of synthetic data to evaluate its strengths and limitations.
The datasets are produced by different models of varying complexity which were designed to assess different aspects of the method.

@@

\item \textbf{Applicability to real-world scenarios:}
The third model contains 103 dipoles with different dipole moment magnitudes, having inclinations and declinations clustered around two stable directions.
The magnetic field generated from these sources is corrupted by both low and high-frequency noise.
The complexity of this synthetic data seeks to more faithfully emulate real magnetic microscopy data.

\item \textbf{Applicability to different grain concentration scenarios:}
The fourth model assesses the algorithm's capability to detect magnetic sources in samples with different magnetic particle concentrations. Several scenario tests were conducted, with progressively increasing magnetic grain concentrations. In parallel with this approach, two dipolar sources were gradually separated, also at different sensor heights, until the algorithm could effectively distinguish the signals from both sources. This additional modeling provides insights into the algorithm's performance in scenarios where magnetic sources are closely spaced, complementing the modeling with higher particle density.

\end{enumerate}

@@

The positions of each source obtained with Euler deconvolution were then used as input for the dipole moment inversion.
The estimated dipole moment vectors are shown in Figure~\ref{fig_synthetic_simple_results}.
Figure~\ref{fig_synthetic_simple_results}a shows the estimated dipole moment, along with the corresponding true values, as a stereonet for better comparison of the true and estimated vector directions.
Figure~\ref{fig_synthetic_simple_results}b shows the estimated dipole moments overlaid on a map of the synthetic $b_z$ to demonstrate the ability of our method to estimate a spatial distribution of dipole moments.
For comparison, Table~\ref{tab_synthetic_simple_results} also shows the true and estimated dipole moments as well as their corresponding standard deviations obtained using Equations~\ref{eq_chi_square}-\ref{eq_variances}.

@@

To simulate particles with non-dipolar magnetization components, we specified a spherical volume with a radius of approximately \qty{1}{\um}.
Within this volume, we added 200 dipolar particles with varying dipole moment directions and amplitudes.
These directions were generated randomly, following a normal distribution centered on the direction $D=\ang{90}$ and $I=\ang{0}$ and with a standard deviation of \ang{10}.
The dipole moment amplitudes were sampled by a normal distribution with mean $\qty{e-14}{\ampere\m\squared}$ and standard deviation $\qty{5e-14}{\ampere\m\squared}$.
We also randomly generated spatial positions around a central position of $x_c=\qty{25}{\um}$, $y_c=\qty{25}{\um}$, and $z_c=\qty{-1}{\um}$ with a standard deviation of \qty{0.65}{\um}.
The bulk magnetization of the non-dipolar particle is the vector sum of the dipole moments of each dipole.
The centroid of the spherical volume is defined by the average of the $x_c$, $y_c$, and $z_c$ positions of each dipole.
The synthetic vertical magnetic field $b_z$ of the non-dipolar source was calculated on regular grids with \qty{0.5}{\um} spacing.
Each grid was located at a different sensor-sample distance $z$, varying between \qty{1}{\um} and \qty{10}{\um} in \qty{0.5}{\um} increments.
The synthetic data were corrupted by pseudo-random Gaussian noise of mean \qty{0}{\nano\tesla} and a standard deviation of \qty{20}{\nano\tesla}.
For each data grid generated, we performed Euler deconvolution to estimate the source position and subsequently inverted the data for the dipole moment.
At each step, we recorded the inversion $R^2$ value and the differences between the estimated position and dipole moment and the true centroid and bulk magnetization of the spherical particle.

@@

We replicated the particle generation procedure 100 times with variations in the randomness of magnetization and point particle positions. Subsequently, the Cartesian position of the spherical particle was estimated using ED, and the magnetization parameters were determined by inverting the magnetic field data. The effectiveness of position recovery was measured by comparing the centroid of the spherical volume with the solution of the Euler deconvolution for $x_c$, $y_c$, and $z_c$ (Figure~\ref{non-dipolarity-synthetic-data-positioning}a, b, and c, respectively). As shown in Figure~\ref{non-dipolarity-synthetic-data-positioning}, ED estimates the positions of particles with more complex magnetization well, especially for sensor-source distances greater than 5 µm, where the median of the distributions is centered at zero. However, when the sensor-source distance is small enough to emphasize the contribution from higher-order components, a lower degree of dipolarity is observed as expressed by lower values of $R^2$ (Figure~\ref{non-dipolarity-synthetic-data-inversion}a). Consequently, in these cases, there are higher errors in the estimation of the direction of magnetization (Figure~\ref{non-dipolarity-synthetic-data-inversion}b) and in the magnetic moment's intensity (Figure~\ref{non-dipolarity-synthetic-data-inversion}c), which were expected since the inversion considers only the dipolar component. However, dipolarity significantly increases as the sensor-to-sample distance increases, while errors in direction and magnetic moment decrease significantly (Figure~\ref{non-dipolarity-synthetic-data-inversion}).

@@

For greater fidelity to real samples, the magnetic sources are modeled as dipoles with different depths and magnetic moment intensities.
The depths vary randomly between 1 and $\qty{20}{\um}$, while the dipole moment intensities range randomly from $10^{-11}$ to $\qty{e-14}{\ampere\m\squared}$.
The NRM found in real ferromagnetic particles varies individually but averages out to the inducing field direction.
To simulate this behavior in our synthetic data, we sample the source dipole moment directions from two pseudo-random Gaussian distributions.
The first group of sources ($M = 70$) are sampled from a distribution  with mean of $D = \ang{0}$ and $I = \ang{0}$ and standard deviation of $\ang{10}$.
The second group of sources ($M = 30$) are sampled from a distribution with mean of $D = \ang{180}$ and $I = \ang{0}$, also with standard deviation of $\ang{10}$. We also manually added 3 sources with higher dipole moments ($5 \times 10^{-11}$) to further simulate the complexity observed in real data measurements. The noise-corrupted synthetic $b_z$ data are shown in Figure~\ref{complex-synthetic-data}a.

@@

We then followed the same processing steps as for the simple synthetic: upward continuation (Figure~\ref{complex-synthetic-data}b),
TGA calculation (Figure~\ref{complex-synthetic-data}c), contrast stretching (Figure~\ref{complex-synthetic-data}d), blob detection (Figure~\ref{complex-synthetic-data}e), and Euler Deconvolution (Figure~\ref{complex-synthetic-data}e).
The Euler deconvolution and dipole moment inversion took approximately 1.5 seconds to compute, while
the source detection algorithm took approximately 8 seconds, on a laptop with an Intel(R) Core(TM) i7-2670QM CPU 2.20GHz with 16 GB of DDR3L 1333 MHz RAM.

A total of 99 sources out of the original 103 were successfully detected. Four sources remained unidentified (black stars in Figure~\ref{complex-synthetic-data}f) due to specific challenges, that arose from two primary factors:

\begin{enumerate}
    \item Three of these sources either overlapped with or were in close proximity to other neighboring sources, making it inherently difficult to distinguish and isolate them.
    \item The identification of one of these sources was impeded by its considerable depth, which led to substantial signal attenuation, further exacerbating the difficulties associated with source recognition.
\end{enumerate}

\begin{figure}[tb!]
\centering
\includegraphics[width=0.75\linewidth]{figures/complex-synthetic-comparison.png}
\caption{
The validation of the result obtained with the inversion was calculated for each individual particle based on the error between the real parameters modeled and their respective recovered values, being (a) the direction and (b) intensity of the magnetic moment, in addition to the $R^2$ score (c) obtained by comparing the forward model and the actual data. The depth and radius of the magnetic sources are also important factors that influence the final result, therefore, these data are given in the form of cross plots, with the vertical bar represented by the depth (1 - \qty{20}{\um}) and the horizontal bar by the dipole moment amplitude ($10^{-11}$ to $\qty{e-14}{\ampere\m\squared}$).
}
\label{complex-synthetic-comparison}
\end{figure}

The spatial resolution of the inversion results is one of the key advantages of magnetic microscopy over the classic techniques of paleomagnetism.
Therefore, we present the inversion results spatially so that we can evaluate any patterns in their distribution.
Figure~\ref{complex-synthetic-comparison} shows the spatial locations of the 99 sources that were identified and the differences between the estimated dipole moments and the true values.
The depth and dipole moment amplitude of each source are represented by horizontal and vertical bars, respectively.
These two variables are useful proxies for the strength and spatial extent of the signal of each source,
which can tell us about the limitations in strengths and sizes of the source signal that the technique is able to correctly invert.
Figure~\ref{complex-synthetic-comparison}a shows the absolute value of the angular difference between the true and the estimated dipole moment vectors.
Figure~\ref{complex-synthetic-comparison}b shows the percentage difference between the true and estimated dipole moment magnitudes ($|100 \left( m_{true} - m_{estimated}\right) ~/~ m_{true}|$).
Figure~\ref{complex-synthetic-comparison}c shows the $R^2$ coefficient (Equation~\ref{eq_r2}), which is equivalent to the non-dipolarity parameter of \citet{Fu2020} and represents how well the dipolar model is able to explain the observed data of each source.
Figure~\ref{complex-synthetic-comparison}d shows the SNR (Equation~\ref{eq_snr}), which expresses the power of the observed data over that of the inversion residuals.
High SNR values correspond to small inversion residuals which indicate that a dipolar model was able to explain the observed data.
Consequently, the variation of SNR values is similar to that of the $R^2$ coefficient.
Figure~\ref{complex-synthetic-comparison} shows that large errors in the estimated dipole moment direction are associated with low values of $R^2$ and SNR.
Conversely, the correlation between errors in dipole moment magnitude and $R^2$ and SNR is less pronounced, with some poor magnitude estimates being associated with $R^2$ and SNR indicating a reasonable fit by a dipolar model.
It is also noticeable that the majority of cases where the direction misfit is given by deep and low-amplitude sources that are close to shallower or higher-amplitude sources.
These results indicated that $R^2$ and SNR can be used as selection criteria to discard sources with likely high errors in the estimated dipole moment.
